---
title: Exploratory Data Analysis
format: typst
---

### Question 1

What district had the closest race between a Democrat and Republican?

### Answer 1

```{r}
svprec1|>
  group_by(CDDIST)|>
  summarize(sumCNGREP01 = sum(CNGREP01),
            sumCNGDEM01 = sum(CNGDEM01),
            sumTotal = sum(cngtotal))|>
  mutate(rep_percent = sumCNGREP01/sumTotal,
         dem_percent = sumCNGDEM01/sumTotal,
         diff = abs(rep_percent-dem_percent))|>
  arrange(diff)|>
  select(sumCNGDEM01,sumCNGREP01, CDDIST, diff)|>
  slice(1)
  
```

Congressional district 45 had the smallest margin of victory, with Democrats
winning the district by .2 percent. 

### Question 2

What district had the most lopsided race?

### Answer 2

```{r}
svprec1|>
  group_by(CDDIST)|>
  summarize(sumCNGREP01 = sum(CNGREP01),
            sumCNGDEM01 = sum(CNGDEM01),
            sumTotal = sum(cngtotal))|>
  mutate(rep_percent = sumCNGREP01/sumTotal,
         dem_percent = sumCNGDEM01/sumTotal,
         diff = abs(rep_percent-dem_percent))|>
  arrange(desc(diff))|>
  select(sumCNGDEM01,sumCNGREP01, CDDIST, diff)|>
  slice(1)
```

Congressional district 11 had the most lopsided race

### Question 3

How many districts featured races between candidates from the same party.

### Answer 3

```{r}
svprec|>
  group_by(CDDIST)|>
  summarize(sumCNGREP01 = sum(CNGREP01),
            sumCNGDEM01 = sum(CNGDEM01),
            sumCNGDEM02 = sum(CNGDEM02),
            sumCNGREP02 = sum(CNGREP02)) |>
  filter((sumCNGDEM01 > 0 & sumCNGDEM02 > 0) | 
           (sumCNGREP01 > 0 & sumCNGREP02 > 0))
```





